Bayesian Sparsification of $\cplx$-valued networks

### 2019-11-06 paper/paper.tex
minor retouching of ARD dropout

The posterior itself is computationally intractable, save for the relatively simple cases
\cite{citation_needed}, and therefore exact inference is traded for tractability and scalability
of Variational Inference. Instead of deriving $p(\omega \mid D)$ using the rule, this approach
recasts posterior inference into variational optimization problem
\begin{equation} \label{eq:variational-progam}
  q_*
    \in \arg \min_{\theta} \rho\bigl(
      q_\theta(\omega), p(\omega \mid D)
    \bigr)
    \,,
\end{equation}
over a tractable parametric family of distributions on the parameter space $
  \mathcal{Q} = \{Q_\theta(d\omega) \colon \theta \in \Theta\}
$. The class $\mathcal{Q}$ is picked so that its member have tractable densities $
  q_\theta(\omega) d\omega
$, can be sampled from and possess tractable $\log$-derivatives $
  \omega \mapsto \nabla_\theta \log q_\theta(\omega)
$, \cite{williams_simple_1992}, or can be represented as differentiable transformations
of non-parametric noise, \cite{kingma_auto-encoding_2014,figurnov_implicit_2019}.
% (reparameterization trick, pathwise gradieent) $q_{\theta}(d\omega)$ is a push-forward
% of $p(d\varepsilon)$ by a differentiable map $g(\varepsilon; \theta)$.
% % http://stillbreeze.github.io/REINFORCE-vs-Reparameterization-trick/#fn:1

@@

Stochastic Gradient Variational Bayes (SGVB), proposed by \cite{kingma_auto-encoding_2014},
is an differentiable unbiased Monte-Carlo estimator of \eqref{eq:elbo_general}, which makes
it possible to employ stochastic gradient-based optimization methods. If $
  \omega \sim q_{\theta}
$ is equivalent in distribution to $
  \omega = g(\varepsilon; \theta)
$ for some non-parametric random variable $
    \varepsilon \sim p_\varepsilon
$ and differentiable $
  g(\varepsilon; \theta)
$, then the SGVB is
\begin{equation} \label{eq:sgvb}
  \widetilde{\mathcal{L}}(\theta; \phi, \lambda)
    = - KL(q_{\theta}(\omega) \| \pi_{\lambda}(\omega))
      % + \frac{N}{\lvert B \rvert} \sum_{i\in B}
        % \log p(z_i \mid \omega_i)b
      + \frac1{L} \sum_{l=1}^L
        \log p_{\phi}(D \mid \omega_{il})
          \Big\vert_{\omega_{l} = g(\varepsilon_{l}; \theta)}
      % =
      % - KL(q_{\theta}(\omega) \| \pi(\omega))
      % + N \mathbb{E}_{B \sim \{D\}_M \otimes q_\theta^M}
        % \hat{\mathbb{E}}_{z,\omega \sim B}
          % \log p(z \mid \omega)
    \,, 
\end{equation}
where $
  (\varepsilon_{l})_{l=1}^L
$ is an iid sample from $p_\varepsilon$. Intractable KL-divergence terms can be also
replaced by differentiable unbiased Monte-Carlo estimators, \cite{kingma_auto-encoding_2014}.
%
For the purposes of this study, we assume that the observed data is independent, and
use minibatch version of \eqref{eq:sgvb} with $L=1$:
\begin{equation} \label{eq:elbo}
  \widetilde{\mathcal{L}}(\theta; \phi, \lambda)
    = - KL(q_{\theta}(\omega) \| \pi_{\lambda}(\omega))
      + \frac{N}{M} \sum_{i=1}^M
        \log p_{\phi}(z_i \mid g(\varepsilon_i; \theta))
    \,,
\end{equation}
for a random subsample $(z_i)_{i=1}^M$ from $D$ and iid $
  (\varepsilon_i)_{i=1}^M \sim p_\varepsilon
$.

@@

Flexibility in dependence of $q_\theta$ on $\alpha$ allows structured dropout with shared
$\alpha$ across groups of parameters $\omega$. \cite{molchanov_variational_2017} study
Variational Dropout from a model sparsification perspective, optimizing $\alpha$ for each
individual parameter. Another variant of Bayesian dropout, called Automatic Relevance
Determination, is developed in \cite{kharitonov_variational_2018}. The key difference from
Variational Dropout is the choice of prior: the log-uniform factor $\pi(\omega_{ij})$ is
replaced by a proper Gaussian prior $
  \mathcal{N}(0, \tau^{-1}_{ij})
$ with learnable precision $\tau_{ij} > 0$. This is known as the Empirical Bayes method,
\cite{citation_needed}, which fits a prior distribution while performing variational inference.
The optimal precision parameter is found in the closed form by minimizing the KL-divergence
term holding variational parameters $\theta$ fixed.

>>>>

The posterior itself is computationally intractable, save for the relatively simple cases
\cite{citation_needed}, and therefore exact inference is traded for tractability and scalability
of Variational Inference. Instead of deriving $p(\omega \mid D)$ using the rule, this approach
recasts posterior inference into variational optimization problem
\begin{equation} \label{eq:variational-progam}
  q_*
    \in \arg \min_{\theta} \rho\bigl(
      q_\theta(\omega), p(\omega \mid D)
    \bigr)
    \,,
\end{equation}
over a tractable parametric family of distributions on the parameter space $
  \mathcal{Q} = \{Q_\theta(d\omega) \colon \theta \in \Theta\}
$, $\theta$ -- generic variational parameter. The class $\mathcal{Q}$ is picked so that
its member have tractable densities $
  q_\theta(\omega) d\omega
$, can be sampled from and possess tractable $\log$-derivatives $
  \omega \mapsto \nabla_\theta \log q_\theta(\omega)
$, \cite{williams_simple_1992}, or can be represented as differentiable transformations
of non-parametric noise, \cite{kingma_auto-encoding_2014,figurnov_implicit_2019}.
% (reparameterization trick, pathwise gradieent) $q_{\theta}(d\omega)$ is a push-forward
% of $p(d\varepsilon)$ by a differentiable map $g(\varepsilon; \theta)$.
% % http://stillbreeze.github.io/REINFORCE-vs-Reparameterization-trick/#fn:1

@@

\textit{Stochastic Gradient Variational Bayes} (SGVB), proposed by \cite{kingma_auto-encoding_2014},
is an differentiable unbiased Monte-Carlo estimator of \eqref{eq:elbo_general}, which makes
it possible to employ stochastic gradient-based optimization methods. If $
  \omega \sim q_{\theta}
$ is equivalent in distribution to $
  \omega = g(\varepsilon; \theta)
$ for some non-parametric random variable $
    \varepsilon \sim p_\varepsilon
$ and differentiable $
  g(\varepsilon; \theta)
$, then the SGVB is
\begin{equation} \label{eq:sgvb}
  \widetilde{\mathcal{L}}(\theta; \phi, \lambda)
    = - KL(q_{\theta}(\omega) \| \pi_{\lambda}(\omega))
      % + \frac{N}{\lvert B \rvert} \sum_{i\in B}
        % \log p(z_i \mid \omega_i)b
      + \frac1{L} \sum_{l=1}^L
        \log p_{\phi}(D \mid \omega_{il})
          \Big\vert_{\omega_{l} = g(\varepsilon_{l}; \theta)}
      % =
      % - KL(q_{\theta}(\omega) \| \pi(\omega))
      % + N \mathbb{E}_{B \sim \{D\}_M \otimes q_\theta^M}
        % \hat{\mathbb{E}}_{z,\omega \sim B}
          % \log p(z \mid \omega)
    \,, 
\end{equation}
where $
  (\varepsilon_{l})_{l=1}^L
$ is an iid sample from $p_\varepsilon$. Intractable KL-divergence terms can be also
replaced by differentiable unbiased Monte-Carlo estimators, \cite{kingma_auto-encoding_2014}.
%
For the purposes of this study, we assume that the observed data is independent, and
use minibatch version of \eqref{eq:sgvb} with $L=1$:
\begin{equation} \label{eq:elbo}
  \widetilde{\mathcal{L}}(\theta; \phi, \lambda)
    = - KL(q_{\theta}(\omega) \| \pi_{\lambda}(\omega))
      + \frac{N}{M} \sum_{i=1}^M
        \log p_{\phi}(z_i \mid g(\varepsilon_i; \theta))
    \,,
\end{equation}
for a random subsample $(z_i)_{i=1}^M$ from $D$ and iid $
  (\varepsilon_i)_{i=1}^M \sim p_\varepsilon
$.

@@

Flexibility in dependence of $q_\theta$ on $\alpha$ allows structured dropout with shared
$\alpha$ across groups of parameters $\omega$. \cite{molchanov_variational_2017} study
Variational Dropout from a model sparsification perspective, optimizing $\alpha$ for each
individual parameter. Another variant of Bayesian dropout, called \textit{Automatic Relevance
Determination}, is developed in \cite{kharitonov_variational_2018}. The key difference is
that the log-uniform factor $\pi(\omega_{ij})$ is replaced by a proper Gaussian prior $
  \mathcal{N}(0, \tau^{-1}_{ij})
$ with learnable precision $\tau_{ij} > 0$. This method, known as \textit{Empirical Bayes}
\cite{citation_needed}, fits a prior distribution while performing variational inference.
Maximizing \eqref{eq:elbo} over $\lambda$, holding other parameters fixed, yields $
  \tau^* = {(\mu^2 + \sigma^2)}^{-1}
$, whence
\begin{equation}  \label{eq:ard-kl-div-real}
  KL(q_\theta(\omega) \| \pi(\omega))
    = \log{\bigl(1 + \tfrac1{\alpha} \bigr)}
    \,.
\end{equation}
